Binary Biases \& Toy Transit Surveys

### 2017-09-22 memos/workdir/toy_analytic_surveys.tex
general cleanup. logg subgiant check. read papers.
renamed things

parsed Furlan et al (2017)'s summary of all the high resolution imaging
following paper.

added the "logg subgiant bias check": Ji Wang claimed in his ~2015 HJ
occ rate paper that like ~50% of the Kepler stars (or something of
equally large number) were subgiants. This was based on an erroneous
extrapolation from Bastien et al's result which showed that for BRIGHT
Kepler stars (where the subgiant contamination is greatest because of a
Malmquist bias). For most of the sample, from 13<Kp<16, it's not such a
big concern.

also wro

\shortauthors{LGB, KM, JNW}
\shorttitle{Toy transit surveys}

@@

\title{ Toy Transit Surveys }

@@

Finally, we perform a synthetic transit survey analogous to {\it Kepler}. We 
show that if we assume every star in this {\it Kepler} survey analog is single, 
we underestimate occurrence rates for a given planet radius and period by 
a factor of $\approx 1.4$, or a percentage error relative to the true 
occurrence rate $\Gamma$ of $\delta\Gamma=-28\%$.
This is relevant for measuring accurate values of the rate at which Earth-like 
planets exist around Sun-like stars.

@@

Ah, transiting planets! We learn so much by studying them. But how much can we 
actually learn, and how much is messed up by binarity?

@@

A more straight-forward effect of binarity on transit surveys is that any star 
which is unknowingly a double introduces extra stars about which planets might 
be detectable.
However, this apparent increase in the number of searchable stars is mitigated 
by dilution: the observed transit depths in such systems are reduced by 
contaminating flux from the binary companion. This dilution is especially 
drastic if the planet orbits the secondary, in which case dilution impacts 
detectability the most.

A related effect to dilution is misinterpretation of transit depths. If the 
planet to star radius ratio is directly taken as the square-root of 
the observed transit depth, derived planetary radii in binary systems will 
systematically too small.
Ciardi et al [2015] explored this issue in depth. They showed that for the case 
of \kepler, assuming host stars to always be single will on 
average underestimate radii by a multiplicative factor of $\approx 1.5$.

Finally, binarity may affect transit surveys through actual astrophysics.
For the relatively rare case [Duquennoy \& Mayor 1991, Raghavan et al 2010] of 
close binaries, dynamical stability imposes a limited range of periods in which 
planets may orbit [e.g., Holman \& Weigert 
1999, Li et al 2017], the effects of which have been observed [Kraus et al 
2016].
Also, the formation rates of planets, and thus their intrinsic occurrence 
rates, may be influenced by properties of circumprimary and/or circumbinary 
disks.
Although the observations are still ongoing, early results seem to indicate 
that once binary companions are $\gtrsim 100\,{\rm AU}$ apart, binarity does 
not decrease planet occurrence rates [CITE], and may even increase them for 
giant planets [Ngo et al 2017].

@@

We can attempt an answer with Monte Carlo simulations of something like the 
\kepler\ field. I say ``something like'' because the binarity properties of the 
\kepler\ field (or more specifically the stars selected in the KIC) are not 
known.
\todo[inline]{but it might be better to do something directly on the KIC's 
selected target stars, like what Ciardi did! The following is a good example, 
but at most it can ``suggest'' that a similar level of error has been made from 
the \kepler\ survey.}

>>>>

\shortauthors{LGB, KM, JNW}
\shorttitle{Binary Biases}

@@

\title{ Binary Biases \& Toy Transit Surveys }

@@

Finally, we perform a synthetic transit survey analogous to {\it Kepler}. We 
show that if we assume every star in this {\it Kepler} survey analog is single, 
we underestimate occurrence rates for a given planet radius and period by 
a factor of $\approx 1.4$, or a percentage error relative to the true 
occurrence rate $\Gamma_{\rm true}$ of $\delta\Gamma_{\rm 
true}=-28\%$. This is relevant for measuring accurate values of the rate at 
which Earth-like planets exist around Sun-like stars.

@@

@@

A more straight-forward effect of binarity on transit surveys is that any star 
which is unknowingly a double introduces extra stars about which planets might 
be detectable.
However, this apparent increase in the number of searchable stars is mitigated 
by dilution: the observed transit depths in such systems are reduced by 
contaminating flux from the binary companion. This dilution is especially 
drastic if the planet orbits the secondary, in which case dilution impacts 
detectability the most.
Note that dilution can come from visually associated companions as well as 
gravitationally bound companions.

A consequence of dilution is misinterpretation of transit depths. If the 
planet to star radius ratio is directly taken as the square-root of 
the observed transit depth, derived planetary radii in binary systems will 
systematically too small.
Ciardi et al [2015] explored this issue in depth. They claimed that for the 
case of \kepler, if ones assumes host stars are always single, one will
underestimate radii by an average multiplicative factor of $\approx 1.5$ 
(presumably, compared to the case of even weights for whether the 
primary/secondary gets a given planet).

Finally, binarity may affect transit surveys through actual astrophysical 
processes.
Duquennoy \& Mayor [1991], and alter Raghavan et al [2010] showed that over 
half of solar-like are separated by less than 100\,AU.
One immediate consequence on circumprimary and circumsecondary architectures is 
that dynamical stability limits begin to carve away in realms of planet space 
where planets might otherwise be detectable [e.g., Holman \& Weigert 1999, Li 
et al 2017].
The formation efficiency of various types of planetary systems may also 
be different in binary systems compared to single-star systems.
For instance, the properties of the circumstellar and circumbinary disks 
influence the dynamics of planetesimals [Rafikov, Silsbee];

 the effects of which have been observed [Kraus et al 
2016].
Also, the formation rates of planets, and thus their intrinsic occurrence 
rates, may be influenced by properties of circumprimary and/or circumbinary 
disks.
Although the observations are still ongoing, early results seem to indicate 
that once binary companions are $\gtrsim 100\,{\rm AU}$ apart, binarity does 
not decrease planet occurrence rates [CITE], and may even increase them for 
giant planets [Ngo et al 2017].

@@

We can attempt an answer with Monte Carlo simulations of something like the 
\kepler\ field. I say ``something like'' because the binarity properties of the 
\kepler\ field (or more specifically the stars selected in the KIC) are not 
known.
\todo[inline]{but it might be better to do something directly on the KIC's 
selected target stars, like what Ciardi did! The following is a good example, 
but at most it can ``suggest'' that a similar level of error has been made from 
the \kepler\ survey.

But it has MAJOR benefits compared to the GIC method.
It eliminates the need to simulate the biases of the KIC (since they're already 
baked in). The process of ``adding'' binaries to the KIC is the closest thing 
to REALITY, since in reality the KIC is missing binary companion information.
It BAKES IN all the real biases that come with mis-estimating stellar 
radii/masses when they are actually binaries (because it's already been done).
It also has literature precedent in Ciardi's work, so you don't need to cook up 
a crazy new procedure -- can just take his with the flaws baked in.

So this method is best if we care about what errors were made in Kepler/the KIC.
It's not as nice if we want something ``general''.
}
